Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-14 csf-2017/paper.tex
Some work on the examples section and discussion.

\textbf{Invoking the adversary} Slightly more complex is the case where we
invoke the adversary. As above, we clear all the registers, except those that
contain arguments the adversary is intended to have access to. We leave a copy
of the stack pointer in $r_\stk$, but only after we have used
the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
we are not using. In one of the registers, we also provide a return pointer,
which we will say more about below.

@@

\textbf{Stack management discipline}
The precautions the stack discipline must take have now been introduced and motivated. Summarized, the stack discipline is:
\begin{description}[font=\normalfont\itshape]
\item[At program startup] a local $\rwlx$-capability (the stack pointer) resides in register $r_\stk$. The system does not hand out global write-local capabilities.
\item[Before returning to the adversary] clear the registers (with words we do not want to pass to the adversary) and erase any part of the stack we have used.
\item[Before invoking the adversary] push an activation record to the stack, create a local $\entry$-capability for the restore instructions of the activation record (the return pointer) and leave it in register $r_0$, leave a copy of the stack capability restricted to the unused part in $r_\stk$, erase the part of the stack we release control over, and clear the non-argument registers.
\item[Before invoking an adversary call] make sure it is a global capability.
\item[When having called by an adversary] make sure the alleged stack pointer has permission $\rwlx$.
\end{description}
\textbf{Reusable macro instructions} 
We define a number of reusable macros. Some of them capture the stack
management discipline and others make the examples in
Section~\ref{sec:examples} readable. The macros are:
\texttt{fetch $r$ $\var{name}$} fetches the capability related to
$\var{name}$ from the linking table and stores it in register
$r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register
$r_\stk$.  \texttt{prepstk $r$} is used when a callback is invoked by
the adversary and prepares the received stack pointer by checking that
it has permission $\rwlx$ and making it conform to our stack
discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps
to the capability in register $r$ in the manner described above. That
is, it pushes local state (the contents of registers
$\overline{r_{\var{priv}}}$), stack pointer and return pointer to the
stack, sets up the necessary on stack return code and creates a return
pointer, restricts the stack pointer, clears the unused part of the
stack, clears the necessary registers and jumps to $r$. Upon return,
the private state is restored. \texttt{mclear $r$} clears all the
memory the capability in register $r$ has authority
over. \texttt{rclear $\var{regSet}$} clears all the registers in
$\var{regSet}$. \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls
  $\overline{(x_i,r_i)}$ $r$} allocates a closure pair where the
capability in register $r$ is used to point to the program and a new
environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an
implicit fetch happens when an instruction refers to $x_i$.

@@

With the logical relation defined, we can now state a strong theorem
about the safety properties of the capability machine:
any capability that only grants reasonable authority is capability
safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
untrusted programs: if the authority granted by
the capabilty for the untrustred program is reasonable, then it is safe to
execute. Intuitively, this is because our logical relation captures the
safe behaviours of the capability machine that do not allow you to
produce something \emph{unsafe} from something safe. 
The Fundamental Theorem also serves as a sanity check of the
logical relation as it guarantees that a large class of capabilities
actually inhabits it.

@@

\subsection{Encapsulation of local state}
The capability machine provides encapsulation of local state. The
programs \texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
\figurename~\ref{fig:prog-f2} demonstrate this.  The two programs
\texttt{f1} and \texttt{f2} are very similar. They both store some
local state, call an untrusted piece of code ($\adv$), and then tests
whether the local state is unchanged. They differ in the way they do
this. Program \texttt{f1} assumes it has access to a stack pointer
and uses the stack to store its local state. It also uses the
stack-based calling convention \texttt{scall} to call the untrusted code. On the
other hand, \texttt{f2} allocates a new piece of memory for its local
state and uses an activation-record based calling convention
(described in the technical appendix) to run the adversarial code.

@@

\begin{figure}[t]
  \centering
\begin{lstlisting}
f2: malloc $r_l$ 1
    store $r_l$ 1
    fetch $r_1$ $\adv$
    call $r_1$($[]$,$[r_l]$)
    assert $r_l$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a calling convention not relyant
    on the stack}
  \label{fig:prog-f2}
\end{figure}

For both programs, we can prove that if they are linked with an
adversary, $\adv$, that is allowed to allocate new memory but
otherwise has no capabilities, then the assertion will never fail when
executing the program (see Lemmas~\ref{lem:correctness-f1}
and~\ref{lem:correctness-f2} below).
% \lars{shouldn't we say already
% here that the adversary is a global capability and that is why,
% intuitively, the local state is preserved ?} \lau{Due to the ``order
% of control'', \texttt{f1} runs first. The lemma is quite restrictive
% on the adversary in the sense that it only has access to malloc (and
% whatever we give et access to). Specifically, it does not have access
% to the stack pointer and because of that we don't need to check it
% first. We could write that the lemmas are limited in their assumptions
% to make them simpler and a more complicated example is coming up.}
The two examples also illustrate the versatility of the logical
relation. The two examples use different calling convention, but the
logical relation is not specific to any calling convention, so we can
use it to reason about both programs.

@@

In order to formulate lemmas about \texttt{f1} and \texttt{f2}, we
need a way to observe whether the assertion fails. To this end, we
assume access to a flag (an address in memory). When the assertion
fails, it sets the flag to $1$ and halts. A slightly simplified
version of the correctness lemma for \texttt{f1} is:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
  where each of the capabilities have an appropriate range of authority and pointer furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the
start configuration is capability safe (in the $\observations$
relation) for a world with a permanent region that requires the
assertion flag to be 0. This is done using an anti-reduction lemma
which states that for a configuration to be capability safe, it is
sufficient to show that it is in it at a later point of the
execution. To reason about \texttt{scall}, it suffices to show that
(1) the configuration just after we jumped is capability safe and (2)
that the configuration just after \texttt{scall} is done cleaning up
is capability safe.  We use the Fundamental Theorem to reason about
the unknown code, but notice that the Fundamental Theorem says nothing
about enter capabilities and that the adversary capability is an enter
capability. Luckily the enter capability becomes a read/execute
capability when we jump and the Fundamental Theorem can be used to
show that this capability is capability safe as a program counter.

@@

\lau{in the following paragraph, we could make a reference to
  \textbf{invoking the adversary} from
  section~\ref{sec:stack-and-return-pointer} where the problem is
  first introduced. In fact, this paragraph is a reiteration of the
  problem in the aformentioned section.}
In \texttt{f3} there are two calls to an adversary and in order for
the assertion in the middle to succeed, they need to be
well-bracketed. If the adversary is able to store the return
pointer from the first call, so it is available for the second call, then it can be used
in a context where \texttt{f3} has $2$ on top of its stack which will cause the
assertion to fail. The return pointer is passed as a local
capability, so the only means the adversary has to store it is the
stack pointer. The stack pointer itself is a local capability,
so the adversary can only store it on the stack. This means that after
the adversary jumps to the return pointer, \texttt{f3} can
revoke it (by erasing it which happens as part of the
\texttt{scall}). In other words, the adversary cannot store the
return pointer for the second call, and it is possible to
prove that the assertion never fails. The correctness lemma for \texttt{f3} is similar to the two previous lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

@@

The final example, in \figurename~\ref{fig:prog-g1}, is a faithful
translation of a trciky example known from the
litterature\cite{XXX}. It consists of two parts, \texttt{g1} and
\texttt{f4}. \texttt{g1} is a closure generator that generates
closures with one variable $x$ set to $0$ in its environment and
\texttt{f4} as the program. The program \texttt{f4} expects two
arguments: a stack and a callback. The stack is necessary to use
\texttt{scall}, but because it comes from an untrusted source, we need
to make sure that it has the correct permission and fits with our
stack dicipline. The callback must be global for the program to work,
we will get back to this point. The program now sets $x$ to $0$ and
calls the callback. When it returns from the callback, it sets $x$ to
$1$ and calls the callback a second time. When it returns from the
callback, it asserts $x$ is $1$ and returns.

@@

To illustrate how subtle this program is, consider one way an
adversary may try to make the assertion fail. In the second callback
an adversary can get to the first callback by calling the closure one
more time. If the adversary can save the return pointer in
the second callback and use it in the first, then the assertion will
fail because $x$ is $0$ before the first callback. The return
pointer cannot be transfered between the callbacks, so this does not
work.

@@

As mentioned, the callback must be global. This is to make sure the
callback is not a piece of code on the stack that restores the
adversaries original stack pointer breaking the encapsulation of
our local stack frame. In the proof of Lemma~\ref{lem:correctness-g1},
this also shows up. When we reason about \texttt{f4}, we revoke all
the local state of the adversary which means that we consider a
private future world. We use the $\execCond{}$ to reason about the
jump to the callback and because the callback is global, it works for
private future worlds. Had the callback been local, then the
$\execCond{}$ would only work for public future worlds, and we would
be stuck in the proof.

In Lemma~\ref{lem:correctness-g1} the order of control has been
inverted compared to the previous lemmas. In this lemma, the adversary
assumes control first with a capability for the closure creator
\texttt{g1}. This means we need to check all arguments are safe
to use and that we clean up before returning in the end. The inversion
of control poses an interesting challenge when it comes to reasoning
about the adversaries local state during the execution of
\texttt{f4}. During the execution of \texttt{f4} and the callbacks, the
adversary should not be able to rely on its local state from before
the call of \texttt{f4}. This is easily done by revoking all the
temporary regions of the world given at the start of \texttt{f4}. When
\texttt{f4} returns, then the adversary is allowed to rely on its
local state again which means that we need to guarantee that the local
state is unchanged. To illustrate how this is done in the proof of
Lemma~\ref{lem:correctness-g1}, consider
\figurename~\ref{fig:worlds-in-pf-g1}. The return pointer given to
\texttt{f4} might be local and it is capability safe in $W_1$, so it
can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed by revoking all
temporary regions and adding a $\iota^\pwl$ region for the stack we
pass away and static temporary regions (a region that accepts exactly
one memory) for our local stack as well as the adversaries local stack.
The given worlds $W_3$ and $W_5$ are public future worlds. None of
these worlds can have masked the temporary regions of $W_1$ with permanent
regions. $W_6$ is constructed by revoking the temporary regions of
$W_5$ and reinstating the temporary regions of $W_1$. This makes $W_6$
a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary.

@@

\begin{itemize}
\item A general well-bracketed control flow result?
  \begin{itemize}
  \item what would that result say?
  \item possible idea: fully abstract compilation from an assembly language with
  a trusted stack to one without
  \item the LR and some of the lemmas already imply well-bracketed control flow, as seen in examples
  \end{itemize}
\item Relation to local parameters in Scala, Algol/Pascal second-class function
  parameters? 
\item Stack clearing realistic?
\item Non-modularity of heap allocation requirement for adversary callbacks
\end{itemize}

It is worth mentioning that local capabilities are quite similar to features
that have been proposed for high-level languages. Specifically,
\citet{osvald_gentrification_2016}'s second-class or local values in Scala are
remarkably similar and we expect they will support similar reasoning patterns.

>>>>

\textbf{Invoking the adversary} Slightly more complex is the case where we
invoke the adversary. As above, we clear all the registers, except those that
contain arguments the adversary is intended to have access to. We leave a copy
of the stack pointer in the appropriate register, but only after we have used
the $\subseg{}{}{}$ to shrink its authority to just the part of the stack that
we are not using. In one of the registers, we also provide a return pointer,
which we will say more about below.

@@

\textbf{Reusable macro instructions} With the high-level approach in mind, we
define the following macro instructions: \texttt{fetch $r$ $\var{name}$} fetches
the capability related to $\var{name}$ from the linking table and stores it in
register $r$. \texttt{push $r$}/\texttt{pop $r$} works as one would expect. All
macros that uses the stack assumes a stack pointer in register $r_\stk$.
\texttt{prepstk $r$} is used when a callback is invoked by the adversary and
prepares the received stack pointer by checking that it has permission $\rwlx$
and making it conform to our stack discipline. \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} Jumps to the
capability in register $r$ in the manner described above. That is, it pushes local state (the
contents of registers $\overline{r_{\var{priv}}}$), stack pointer and return
pointer to the stack, sets up the necessary on stack return code and creates a
return pointer, restricts stack pointer, clears the unused part of the stack,
clears the necessary registers and jumps to $r$. Upon return, the private state
is restored. \texttt{mclear $r$} clears all the memory the capability in
register $r$ has authority over. \texttt{rclear $\var{regSet}$} clears all the
registers in $\var{regSet}$ \texttt{reqglob $r$} checks whether the word in
register $r$ is a global capability. \texttt{crtcls $\overline{(x_i,r_i)}$ $r$}
allocates a closure pair where the capability in register $r$ is used to point
to the program and a new environment is allocated (using malloc) where the contents of
$\overline{r_i}$ is stored. In the program referred to by $r$, an implicit fetch
happens when an instruction refers to $x_i$.

@@

\subsection{Safety of the capability machine}
With the logical relation defined, we can now state the fundamental theorem of
our logical relation: a strong theorem that formalises the guarantees offered by
the capability machine. Essentially, it says that a capability that only grants
safe authority is capability safe as a program counter.
\begin{theorem}[Fundamental Theorem]
  \label{thm:ftlr}
  For all $n$, $\perm$, $\start$, $\addrend$, $\addr$, $\gl$, $W$, if one of the following holds:
  \begin{align*}
      & \bullet
        \begin{aligned}[t]
        &\perm = \exec \text{\footnotesize{ and }}\npair{(\start,\addrend)} \in \readCond{}(\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwx \text{\footnotesize{ and }} \\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W)
      \end{aligned} \\
    & \bullet 
      \begin{aligned}[t]
        &\perm = \rwlx \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \text{\footnotesize{ and }}\\
        &\npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W),
      \end{aligned}
  \end{align*}
  then $\npair{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W)$
\end{theorem}
The Fundamental Theorem allows us to prove capability safety for
untrusted programs: if the authority granted by
the capabilty for the untrustred program is reasonable, then it is safe to
execute. Intuitively, this is because our logical relation captures the
safe behaviours of the capability machine that do not allow you to
produce something \emph{unsafe} from something safe. 
The Fundamental Theorem also serves as a sanity check of the
logical relation as it guarantees that a large class of capabilities
actually inhabits it.

@@

\subsection{Encapsulation of local state}
The capability machine provides encapsulation of local state. The programs
\texttt{f1} in \figurename~\ref{fig:prog-f1} and \texttt{f2}
\figurename~\ref{fig:prog-f2} demonstrate this. They are very similar: both
store some local state, call an untrusted piece of code ($\adv$), and then test
whether the local state is unchanged. They differ in the way they do this.
Program \texttt{f1} assumes it has access to a stack pointer and uses the stack
to store its local state. It also uses the stack-based calling convention
\texttt{scall} to call the untrusted code. On the other hand, \texttt{f2} uses
malloc to allocate memory for its local state and uses an activation-record
based calling convention (described in the technical appendix) to run the
adversarial code.

@@

\begin{figure}[t]
  \centering
\begin{lstlisting}
f2: malloc $r_l$ 1
    store $r_l$ 1
    fetch $r_1$ $\adv$
    call $r_1$($[]$,$[r_l]$)
    assert $r_l$ 1
    halt
\end{lstlisting}
  \caption{Simple program that uses a calling convention not reliant
    on the stack}
  \label{fig:prog-f2}
\end{figure}

For both programs, we can prove that if they are linked with an
adversary, $\adv$, that is allowed to allocate new memory but
otherwise has no capabilities, then the assertion will never fail when
executing the program (see Lemmas~\ref{lem:correctness-f1}
and~\ref{lem:correctness-f2} below).
% \lars{shouldn't we say already
% here that the adversary is a global capability and that is why,
% intuitively, the local state is preserved ?} \lau{Due to the ``order
% of control'', \texttt{f1} runs first. The lemma is quite restrictive
% on the adversary in the sense that it only has access to malloc (and
% whatever we give et access to). Specifically, it does not have access
% to the stack pointer and because of that we don't need to check it
% first. We could write that the lemmas are limited in their assumptions
% to make them simpler and a more complicated example is coming up.}
The two examples also illustrate the versatility of the logical relation.
Because it is not specific to any calling convention, we can use it to reason
about both programs, even though they use different calling conventions.

@@

In order to formulate lemmas about \texttt{f1} and \texttt{f2}, we
need a way to observe whether the assertion fails. To this end, we
assume access to a flag (an address in memory). When the assertion
fails, it sets the flag to $1$ and halts. A slightly simplified
version of the correctness lemma for \texttt{f1} is:
\begin{lemma}
  \label{lem:correctness-f1}
  Let
\[
    \begin{array}{r@{}lr@{}l}
    c_{\var{adv}} & \;\defeq ((\entry,\glob),\dots) & c_{\var{stk}} & \;\defeq ((\rwlx,\local),\dots)\\
    c_{f1} & \;\defeq ((\rwx,\glob),\dots) & c_\link &\; \defeq ((\readonly,\glob),\dots)\\
    c_\malloc &\; \defeq ((\entry,\glob),\dots) & \reg& \; \in \Regs \\
    m &  \multicolumn{3}{@{}l}{\;\defeq \ms_{f1} \uplus \ms_\flag \uplus \ms_{\var{link}} \uplus \hs_\adv \uplus \ms_{\malloc} \uplus} \\
      & \multicolumn{3}{@{}l}{\phantom{\;\defeq \;}  \ms_{\var{stk}} \uplus \ms_{\var{frame}}} \\
    \end{array}
\]
  where each of the capabilities have an appropriate range of authority and pointer. Furthermore
  \begin{itemize}
  \item $\ms_{f1}$ contains $c_\link$, $c_\flag$ and the code of \texttt{f1}
  \item $\ms_\flag(\flag) = 0$
  \item $\ms_{\var{link}}$ contains $c_\adv$ and $c_\malloc$
  \item $\hs_\adv$ contains $c_\link$ and otherwise only instructions.
  \end{itemize}
  If $(\reg\update{\pcreg}{c_{f1}}\update{r_\stk}{c_\stk},m) \step^* (\halted,m')$,
  then $m'(\flag) = 0$
\end{lemma}

To prove Lemma~\ref{lem:correctness-f1}, it suffices to show that the start
configuration is capability safe (in the $\observations$ relation) for a world
with a permanent region that requires the assertion flag to be 0. We prove this
using an anti-reduction lemma by which it suffices to show that the
configuration is safe after some reduction steps. We then use a general lemma
for reasoning about \texttt{scall}, by which it suffices to show that (1) the
configuration that \texttt{scall} will jump to is capability safe and (2) that
the configuration just after \texttt{scall} is done cleaning up is capability
safe. We use the Fundamental Theorem to reason about the unknown code, but
notice that the Fundamental Theorem says nothing about enter capabilities and
that the adversary capability is an enter capability. Luckily the enter
capability becomes $\exec$ after the jump and then, the Fundamental Theorem
applies.

@@

In \texttt{f3} there are two calls to an adversary and in order for the
assertion in the middle to succeed, they need to be well-bracketed. If the
adversary were able to store the return pointer from the first call and invoke
it in the second call, then \texttt{f3} would have $2$ on top of its stack and
the assertion will fail. However, the security measures of our calling
convention explained in Section~\ref{sec:stack-and-return-pointer} prevent this
sort of attack: specifically, the return pointer is local, so it can only be
stored on the stack, but the part of the stack that is accessible to the
adversary is cleared before the second invocation. IThe following lemma shows
that there are also no other attacks that will break well-bracketedness of this
example, i.e. the assertion never fails. It is similar to the two previous
lemmas:
\begin{figure}[t]
  \centering
  \begin{lstlisting}
f3: push 1
    fetch $r_1$ $\adv$
    scall $r_1$($[]$,$[r_1]$)
    pop $r_2$
    assert $r_2$ 1
    push 2
    scall $r_1$($[]$,$[]$)
    halt
\end{lstlisting}
  \caption{Simple program that relies on control-flow integrity.}
  \label{fig:prog-f3}
\end{figure}

@@

The final example, in \figurename~\ref{fig:prog-g1}, is a faithful translation
of a tricky example known from the
literature~\citep{pitts_operational_1998,Dreyer:jfp12}. It consists of two
parts, \texttt{g1} and \texttt{f4}. \texttt{g1} is a closure generator that
generates closures with one variable $x$ set to $0$ in its environment and
\texttt{f4} as the program. \texttt{f4} expects one argument, a callback. The
program sets $x$ to $0$ and calls the callback. When it returns, $\texttt{f4}$
sets $x$ to $1$ and calls the callback a second time. When it returns from the
callback, it asserts that $x$ is $1$ and returns. This example is more
complicated than the previous ones because it involves a closure invoked by the
adversary and an adversary callback invoked by us. As explained in
Section~\ref{sec:stack-and-return-pointer}, this means we need to (1) take care
to properly check the stack pointer received by our closure from the adversary
(it must have the correct permission and fit with our stack discipline) and (2)
require that the adversary callback be global.

@@

To illustrate how subtle this program is, consider how an adversary could try to
make the assertion fail. In the second callback an adversary can get to the
first callback by invoking the closure one more time. If there were any way for
the adversary to transfer the return pointer from the point where it reinvokes
the closure to where the closure reinvokes the callback, then the assertion can
be made to fail. Similarly, if there were any way for the adversary to store a
stack pointer or trick the trusted code into preserving it across an invocation,
the assertion can likely be made to fail too. However, our stack and return
pointer discipline guarantees that there is no way any of this can happen, as we
prove in the following lemma.

@@

As explained in Section~\ref{sec:stack-and-return-pointer}, the callback must be
global, essentially to make sure that it is not allocated on the stack where it
might contain the adversary's original stack pointer. If not, it could break the
encapsulation of our local stack frame. In the proof of
Lemma~\ref{lem:correctness-g1}, this requirement shows up because we invoke the
callback in a world that is only a private future world of the one where we
received the callback, precisely because we have invalidated the local state of
the adversary (particularly its previous stack and return capabilities). The
callback is still valid in this private future world, but only because we know
that it is global.

In Lemma~\ref{lem:correctness-g1} the order of control has been inverted
compared to the previous lemmas. In this lemma, the adversary assumes control
first with a capability for the closure creator \texttt{g1}. This means we need
to check that all arguments are safe to use and that we clean up before
returning in the end. The inversion of control poses an interesting challenge
when it comes to reasoning about the adversary's local state during the
execution of \texttt{f4}. During the execution of \texttt{f4} and the callbacks,
the adversary should not be able to rely on its local state from before the call
of \texttt{f4}. This is easily done by revoking all the temporary regions of the
world given at the start of \texttt{f4}. However, when \texttt{f4} returns, the
adversary is again allowed to rely on its old local state which means that we
need to guarantee that the local state is unchanged. This is important because
the return pointer that \texttt{f4} receives from the adversary may be local,
and the adversary is allowed to allocate the activation record on the stack
(just like we do) so that he can store and recover his old stack pointer after
\texttt{f4} returns.

In Figure~\ref{fig:worlds-in-pf-g1}, we illustrate how we accomodate this in the
proof of Lemma~\ref{lem:correctness-g1} by constructing appropriate worlds for
all the situations where control is passed to the adversary. The potentially
local return pointer received by \texttt{f4} from the adversary is safe in
$W_1$, so it can only be used in public future worlds. Let us see how $W_6$ is
constructed as a public future world. The worlds $W_2$ and $W_4$ are constructed
by revoking all temporary regions and adding a $\iota^\pwl$ region for the stack
we pass away and static temporary regions (a region that accepts exactly one
memory) for our local stack as well as the adversaries local stack. The given
worlds $W_3$ and $W_5$ are public future worlds in which the adversary chooses
to return to us. None of these worlds can have masked the temporary regions of
$W_1$ with permanent regions. $W_6$ is constructed by revoking the temporary
regions of $W_5$ and reinstating the temporary regions of $W_1$. This makes
$W_6$ a public future world of $W_1$. It is therefore safe to use the return
pointer to return to the adversary, since we have restored validity of any local
state he/she might have stashed away.

@@

\textbf{Formulating control flow integrity} The attentive reader may notice
that, while we claim that our stack and return pointer discipline enforces
control flow integrity, we do not prove a general theorem to show that this is
the case. The reason for this is that it is not clear what such a theorem should
look like. Formulations in terms of a control flow graph, like the one by
\cite{abadi_control-flow_2005}, do not take into account temporal properties,
like the well-bracketedness that Example~\texttt{g1} relies on.\footnote{It is a
  bit unfortunate that the general term ``control flow integrity'' is used for a
  property that only captures partial correctness of control flow.} In fact, our
examples show that our logical relation imply a form of control flow integrity
that is stronger than such formulations, although this is not made very
explicit. As future work, we are considering to look at a more explicit and
useful way to formalise well-bracketed control flow. The idea would be to define
a variant of our capability machine with built-in call and return instructions
and well-bracketed control flow built-in to the operational semantics, and then
prove that compiling such code to our machine using our stack and return pointer
discipline is fully abstract~\cite{abadi_protection_1998}, which roughly means
that the lower-level machine does not allow interactions with the compiled code
that do not correspond to an interaction with the original code in the
high-level machine.

\textbf{Performance and the requirement for stack clearing} While performance is
not a goal of this

\textbf{Modularity} Our return pointers and callbacks satisfy the checks we
apply to the ones we get from the adversary (return pointers are allowed to be
local, callbacks must be global), i.e. we do not claim special status and what
we propose is modular.

\textbf{Local capabilities in high-level languages} It is worth mentioning that
local capabilities are quite similar to features that have been proposed for
high-level languages. Specifically, \citet{osvald_gentrification_2016}'s
second-class or local values in Scala are remarkably similar: they are a kind of
values that can be provided to other code for immediate use without allowing it
to store the value in a closure or reference for later use. We believe reasoning
about such values will require very similar reasoning techniques than the ones
we have used for local capabilities.
